---
title: "Effect of Treatment Over Time"
format: html
Author: Muhammad Shakir Nasir 
---

# 2.2 Effect of Treatment Over Time

## 2.2.1 The Data

### Description of the Dataset:

The dataset contains performance measurements recorded at two different time points: before and after treatment. It includes two variables: performance (`perf`) and time (`time`).

### Description of the Measured Variables

| Variable | Description | Unit | Possible values | Expected change |
|----|----|----|----|----|
| perf | Performance measurement | Not specified | Numerical values | May increase after treatment |
| time | Measurement time | Not applicable | Before, After | Not applicable |

## 2.2.2 The Analysis

### Research Question
Does performance differ before and after treatment?
### Descriptive Statistics
```{r}
# Import the dataset
data <- read.csv("DATA/PrePost.csv")

# Display the dataset
data

# Check the structure
str(data)

# Check the number of observations
nrow(data)
```

```{r}
# Count observations in each time condition
table(data$time)

# Descriptive statistics for each condition
by(data$perf, data$time, summary)

# Standard deviation for each condition
by(data$perf, data$time, sd)
```
### Statistical Test

### Data Visualization

### Interpretation of Results

### Conclusion
